Understand whether or not the program has different impacts on the two subgroups, individuals with or without high school degrees. 
- Is there a difference?
- How is there a difference?
  - Multivariate Regression with re78 and incorporating at least 6 predictors, including treatment status -
  
Part 1:
```{r}
require("tableone")
require("MatchIt")
require("Matching")
library(tableone)
library(MatchIt)
library(Matching)
data(lalonde, package = "Matching")

xvars = c("age","educ","black","hisp","married","nodegr","re74","re75", "re78")
table1<- CreateTableOne(vars=xvars, strata="treat", data = lalonde, test=FALSE)
print(table1, smd = TRUE)
```

```{r}
psmodel = glm(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, family = binomial(),data = lalonde)
summary(psmodel)
pscore<-psmodel$fitted.values
```
```{r}
set.seed(931139)
psmatch = Match(Tr = lalonde$treat,M = 1,X = pscore, replace = FALSE, caliper = NULL)
matched = lalonde[unlist(psmatch[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)

```
```{r}
set.seed(931139)
logit = function(p) {log(p)-log(1-p)}
psmatch = Match(Tr = lalonde$treat,M = 1,X = pscore, replace = FALSE, caliper = 0.2)
matched = lalonde[unlist(psmatch[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)

```
```{r}
y_trt = matched$re78[matched$treat == 1]
y_con = matched$re78[matched$treat == 0]
diffy = y_trt - y_con
mean(y_trt) - mean(y_con) 
t.test(diffy)
```

Starting from the very beginning and doing matching among students with, and without high school degrees
```{r}
require("tableone")
require("MatchIt")
require("Matching")
library(tableone)
library(MatchIt)
library(Matching)
data(lalonde, package = "Matching")

wdeg = subset(lalonde[lalonde$nodegr == 0,], select = -c(nodegr))
nodeg = subset(lalonde[lalonde$nodegr == 1,], select = -c(nodegr))

xvars = c("age","educ","black","hisp","married","re74","re75", "re78")
tablewdeg = CreateTableOne(vars=xvars, strata="treat", data = wdeg, test=FALSE)
print(tablewdeg, smd = TRUE)

xvars = c("age","educ","black","hisp","married","re74","re75", "re78")
tablenodeg = CreateTableOne(vars=xvars, strata="treat", data = nodeg, test=FALSE)
print(tablenodeg, smd = TRUE)

```
There might be a problem - the sample of students with degrees isn't very large.
```{r}
psmwdeg = glm(treat ~ age + educ + black + hisp + married + re74 + re75, family = binomial(),data = wdeg)

set.seed(931139)
psmatchwdeg = Match(Tr = wdeg$treat, M = 1, X = psmwdeg$fitted.values, replace = FALSE, caliper = NULL)
matchedwdeg = wdeg[unlist(psmatchwdeg[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","re74","re75")


matchedtab1wdeg = CreateTableOne(vars = xvars, strata ="treat", data = matchedwdeg, test = FALSE)
print(matchedtab1wdeg, smd = TRUE)


psmnodeg = glm(treat ~ age + educ + black + hisp + married + re74 + re75, family = binomial(),data = nodeg)

set.seed(931139)
psmatchnodeg = Match(Tr = nodeg$treat, M = 1, X = psmnodeg$fitted.values, replace = FALSE, caliper = NULL)
matchednodeg = nodeg[unlist(psmatchnodeg[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","re74","re75")


matchedtab1wdeg = CreateTableOne(vars = xvars, strata ="treat", data = matchednodeg, test = FALSE)
print(matchedtab1wdeg, smd = TRUE)
```
Unfortunately more treatment than control

```{r}
trtdgr = matchedwdeg$re78[matchedwdeg$treat == 1] 
condgr = matchedwdeg$re78[matchedwdeg$treat == 0 ]
mean(trtdgr) - mean(condgr)
diffdgr = trtdgr - condgr
t.test(diffdgr, alternative = "greater")

trtnodgr = matchednodeg$re78[matchednodeg$treat == 1] 
connodgr = matchednodeg$re78[matchednodeg$treat == 0]
mean(trtnodgr) - mean(connodgr)
diffnodgr = trtnodgr - connodgr
t.test(diffnodgr, alternative = "greater")
```




Now we have the treatment effect and confidence intervals.
```{r}
trtdgr = matched$re78[matched$treat == 1 & matched$nodegr == 1] 
condgr = matched$re78[matched$treat == 0 & matched$nodegr == 1]
mean(trtdgr) - mean(condgr)
diffdgr = trtdgr - condgr
t.test(diffdgr)

trtnodgr = matched$re78[matched$treat == 1 & matched$nodegr == 0] 
connodgr = matched$re78[matched$treat == 0 & matched$nodegr == 0]
mean(trtnodgr) - mean(connodgr)
diffnodgr = trtnodgr - connodgr
t.test(diffnodgr)
```










```{r}


mean(lalonde$re78[lalonde$treat == 1]) - mean(lalonde$re78[lalonde$treat == 0])

summary(lm(re78 ~ age + educ + black + hisp + married + nodegr + re74 + re75 + treat, data = lalonde))
```

    - **how should we choose these 6 or should we choose them all?**
    - What are the effects and statistical significance of the treatment on re78
    - Estimate confidence intervals (using standard error results or via simulation)
    - **Does regression allow for different effects for different individuals? Explain**
    - Are estimated treatment effects higher for indivudals with or without high school degree?
  - Random forest with same predictors as in 2, and
```{r}
library(randomForest)
rf = randomForest(re78 ~ age + educ + black + hisp + married + nodegr + re74 + re75 + treat, data = lalonde)
rf$importance
varImpPlot(rf, main = "Variables in Importance")
```
  
    - copy-paste variable importance plot into assignment
    - Use random forest to estimate average treatment effect for the units in the treatment group
    - Does your tree produce different effects for different individuals?
    - Are estimated treatment effects higher for indivudals with or without high school degree?
```{r}
sum(lalonde$treat == 1 & lalonde$nodegr == 1)
sum(lalonde$treat == 1 & lalonde$nodegr == 0)
sum(lalonde$treat == 0 & lalonde$nodegr == 0)
sum(lalonde$treat == 0 & lalonde$nodegr == 1)
```
    
    - Specify dataset in the function and don't use $
    - Predict counterfactual potential outcomes for the two new datasets, one with high school grads and one with no high school grads - we have to stratefy over high school graduation. The challenge is that non-high school grads will often have lower earning potential
```{r}
nodegr = subset(lalonde, nodegr == 1)
degr = subset(lalonde, nodegr == 0)
tapply(nodegr$re78, nodegr$treat, summary)
tapply(degr$re78, degr$treat, summary)
```
Expected value involving discrete + continuous?

Estimate the propensity score (using logistic regression/random forest)
```{r}
glm1 = glm(treat ~ age + educ + black + hisp + married + nodegr, data = lalonde, family=binomial())
glm2 = glm(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75 , data = lalonde, family=binomial())
glm1$fitted.values
glm2$fitted.values
```

```{r}
m.out1 = matchit(treat ~ age + educ + black + hisp + married + nodegr, data = lalonde, method = "nearest")
plot(m.out1, type = "jitter")
plot(m.out1, type = "hist")
summary(m.out1)
m.out1optimal = matchit(treat ~ age + educ + black + hisp + married + nodegr, data = lalonde, method = "optimal")
plot(m.out1optimal, type = "jitter")
plot(m.out1optimal, type = "hist")
summary(m.out1optimal)
# m.out1full = matchit(treat ~ age + educ + black + hisp + married + nodegr, data = lalonde, method = "full")
# plot(m.out1full, type = "jitter")
# plot(m.out1full, type = "hist")
# summary(m.out1full)
# m.out1genetic = matchit(treat ~ age + educ + black + hisp + married + nodegr, data = lalonde, method = "genetic")
# plot(m.out1genetic, type = "jitter")
# plot(m.out1genetic, type = "hist")
# summary(m.out1genetic)
m.out1optimal
```

Nearest neighbor algorithm is 
- computationally fast, even for large datasets
- Not invariant ot the intiial order of list
- Not optimal, and can lead to some bad matches

Largely a bias - variance tradeoff issue.

-optmatch or rcbalance

Checking covariate balance using tableone, or smd = standardized means difference
rule of thumb is that <0.1 indicates adequate balance, 0.1-0.2 is not too alarming, and >0.2 indicates serious imbalance



```{r}
m.out2 = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "nearest")
plot(m.out2, type = "jitter")
plot(m.out2, type = "hist")
summary(m.out2)
m.out2optimal = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "optimal")
plot(m.out2optimal, type = "jitter")
plot(m.out2optimal, type = "hist")
summary(m.out2optimal)
m.out2full = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "full")
plot(m.out2full, type = "jitter")
plot(m.out2full, type = "hist")
summary(m.out2full)
m.out2genetic = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "genetic")
plot(m.out2genetic, type = "jitter")
plot(m.out2genetic, type = "hist")
summary(m.out2genetic)
```

needs graph

Trim tails to remove control subjects with propensity score less than minimum in treatment group and treated subjects with propensity score greater than maximum in control group
```{r}
pscore1 = glm(treat ~ age + educ + black + hisp + married + nodegr, data = lalonde, family=binomial())$fitted.values
psmatch = Match(Tr=lalonde$treat, M=1, X=logit(pscore1), replace=FALSE, caliper = 0.1)
matched = lalonde[unlist(psmatch[c("index.treated", "index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75","u74","u75")
matchedtab1 = CreateTableOne(vars = xvars, strata="treat", data= matched, test = FALSE)
print(matchedtab1, smd = TRUE)
```


, and match on logit using nearest neighbor/optimal matching of said propensity score (the propensity score is bounded between 0-1, making many values seem similar, and logit of proensity score is unbounded which stretches the distribution while preserving ranks). using plot


Caliper to ensure no bad matches are accepted - commonly used caliper is 0.2 times standard deviation of logit of propensity score
1. estimate propensity score
2. logit propensity score
3. take standard deviation
4. set caliper to 0.2 times standard deviation
a smaller caliper means less bias but more variance

Average causal effect in dollars

Confidence interval using T-test

Use a paired t-test

Model-based estimates
```{r}
# m.out() = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "optimal")

m.out0 = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "nearest", discard = "hull.control")
z.out0 = zelig(re78 ~ treat + age + educ + black + nodegr + re74 + re75, data = match.data(m.out0), model = "ls")
x.out0 = setx(z.out0, treat = 0)
x1.out0 = setx(z.out0, treat = 1)
s.out0 = sim(z.out0, x = x.out0, x1 = x1.out0)
summary(s.out0)

```

Average treatment effect on the treated
```{r}
m.out1 = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "nearest")
z.out1 = zelig(re78 ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = match.data(m.out1, "control"), model = "ls" )
x.out1 = setx(z.out1, data = match.data(m.out1, "treat"), cond = TRUE)
s.out1 = sim(z.out1, x = x.out1)
summary(s.out1)
```

Average treatment effect overall
```{r}
m.out1 = matchit(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, data = lalonde, method = "nearest")

z.out = zelig(re78 ~ treat + age + educ + black + hisp + married + nodegr + re74 + re75, data = match.data(m.out), model = "ls")
x.out = setx(z.out, treat = 0)
x1.out = setx(z.out, treat = 1)
s.out = Zelig::sim(z.out, x = x.out, x1 = x1.out)
summary(s.out)
# 
# 
# 
# x.out2 = setx(z.out2, data = match.data(m.out1, "control"), cond = TRUE)
# s.out2 = sim(z.out2, x = x.out2)
# 
# ate.all = c(s.out1$qi$att.ev, -s.out2$qi$att.ev)
# 
# mean(ate.all)
# sd(ate.all)
# quantile(ate.all, c(0.025, 0.975))

```



https://dash.harvard.edu/bitstream/handle/1/11130519/King_MatchIt.pdf?sequence%3D1

    - Bootstrap confidence intervals for random forest results
  - Implement a Fischer Exact Test using the sharp null hypothesis of no treatment effectfor any unit, and interpret results. Produce a data visualization that shows to what extent the observed values - re78 is unusual if the null hypothesis is true
  - Is there a way to use FET methodology to establish if treatment is higher for individuals with and without a college degree? Explain
  How to interpret results from an observational study? The fact that it wasn't automatically changed makes it OK?

