---
title: "Causal Inference Assignment"
output: html_notebook
---
Understand whether or not the program has different impacts on the two subgroups, individuals with or without high school degrees. 
- Is there a difference?
- How is there a difference?
  - Multivariate Regression with re78 and incorporating at least 6 predictors, including treatment status -
```{r}
require("tableone")
require("MatchIt")
require("Matching")
library(tableone)
library(MatchIt)
library(Matching)
data(lalonde, package = "Matching")

xvars = c("age","educ","black","hisp","married","nodegr","re74","re75", "re78")
table1<- CreateTableOne(vars=xvars, strata="treat", data = lalonde, test=FALSE)
print(table1, smd = TRUE)
psmodel = glm(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, family = binomial(),data = lalonde)
summary(psmodel)
pscore<-psmodel$fitted.values
range(pscore)

set.seed(931139)
psmatch = Match(Tr = lalonde$treat,M = 1,X = pscore, replace = FALSE, caliper = NULL)
matched = lalonde[unlist(psmatch[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)

set.seed(931139)
logit = function(p) {log(p)-log(1-p)}
psmatch = Match(Tr = lalonde$treat,M = 1,X = pscore, replace = FALSE, caliper = 0.2)
matched = lalonde[unlist(psmatch[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)

y_trt = matched$re78[matched$treat == 1]
y_con = matched$re78[matched$treat == 0]
diffy = y_trt - y_con
mean(y_trt) - mean(y_con) 
t.test(diffy)
```




Part 1
Go to http://users.nber...
Load these data into R. These are the data from the randomized experiment. 
Obtain the point estimate of the treatment effect by running a simple difference in means. 
Obtain the confidence interval by running a univariate linear regrssion (with just he treat variable).

```{r}
# Load the data into R
treated = read.table("http://users.nber.org/~rdehejia/data/nswre74_treated.txt")
control = read.table("http://users.nber.org/~rdehejia/data/nswre74_control.txt")

colnames = c("treat", "age", "educ", "black", "hisp", 
             "married", "nodegr", "re74", "re75", "re78")

names(treated) = colnames
names(control) = colnames

# Obtain point estimate of treatment effect by running a simple difference in means
mean(treated$re78) - mean(control$re78)

# Obtain confidence interval by running a univariate linear regression with (with just the treat variable). Report your results
confint(lm(re78 ~ treat, data = rbind(treated, control)))
```
Seems to be an effect of 1794, confidence interval is 550-3038


Part 2
Now substitute the cps_controls.txt or equivalently the cps_controls for the control group from the randomized epxeriment. Repeat what you did in #1 above.
```{r}
control = read.table("http://users.nber.org/~rdehejia/data/cps_controls.txt")
names(control) = colnames

# Obtain point estimate of treatment effect by running a simple difference in means
mean(treated$re78) - mean(control$re78)

# Obtain confidence interval by running a univariate linear regression with (with just the treat variable). Report your results
confint(lm(re78 ~ treat, data = rbind(treated, control)))
```
Seems to be an effect of -8497, confidence interval is (-9893)-(-7101.87)
  
Part 3
Use the Match() help file code example as a guide, use propensity score matching to produce an estimated treatment effect and confidence interval. 
Report your results.
```{r}
lalonde['u74'] = lalonde$re74 == 0
lalonde['u75'] = lalonde$re75 == 0
glm1 = glm(treat ~ age + I(age^2) + educ + I(educ^2) + black + hisp + married + nodegr + re74 + I(re74^2) + re75 + I(re75^2) + u74 + u75, family = binomial, data = lalonde)

X = glm1$fitted.values
Y = lalonde$re78
Tr = lalonde$treat

rr = Match(Y = Y, Tr = Tr, X = X, M = 1)
summary(rr)
# 
# mb = MatchBalance(treat ~ age + I(age^2) + educ + I(educ^2) + black + hisp + married + nodegr + re74 + I(re74^2) + re75 + I(re75^2) + u74 + u75, data = lalonde, match.out = rr, nboots = 10)
# 

matched = lalonde[unlist(rr[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75", "u74", "u75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)
```



Part 4  
Use the Match() function to run a multivariate matching procedure that uses all the covariates and also includes the estimated propensity scores. 
Report your results.
```{r}
require("Matching")

lalonde = rbind(treated, control)


xvars = c("age","educ","black","hisp","married","nodegr","re74","re75", "re78")
table1<- CreateTableOne(vars=xvars, strata="treat", data = lalonde, test=FALSE)
print(table1, smd = TRUE)
psmodel = glm(treat ~ age + educ + black + hisp + married + nodegr + re74 + re75, family = binomial(),data = lalonde)
summary(psmodel)
pscore<-psmodel$fitted.values
range(pscore)

set.seed(931139)
psmatch = Match(Tr = lalonde$treat,M = 1,X = pscore, replace = FALSE, caliper = NULL)
matched = lalonde[unlist(psmatch[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)

set.seed(931139)
logit = function(p) {log(p)-log(1-p)}
psmatch = Match(Tr = lalonde$treat,M = 1,X = logit(pscore), replace = FALSE, caliper = 0.05)
matched = lalonde[unlist(psmatch[c("index.treated","index.control")]), ]
xvars = c("age","educ","black","hisp","married","nodegr","re74","re75")

#get standardized differences
matchedtab1 = CreateTableOne(vars = xvars, strata ="treat", data = matched, test = FALSE)
print(matchedtab1, smd = TRUE)

y_trt = matched$re78[matched$treat == 1]
y_con = matched$re78[matched$treat == 0]
diffy = y_trt - y_con
mean(y_trt) - mean(y_con) 
t.test(diffy)


# match = Match(Y = df$re78, Tr = df$treat, X = psm$fitted.values, M = 1)
# summary(match)
# MatchBalance(treat ~ age + educ + black + hisp + married + nodegr + re74  + re75 , data = df, match.out = match, nboots = 50)

# Sicars
# mean(df$re78[match$index.treated]) - mean(df$re78[match$index.control])
# Nik
# coef(lm(Y ~ Tr, data = match$mdata))[2]

# Confidence interval method 1: Using 1.96
# c(match$est - 1.96 * match$se, match$est + 1.96 * match$se)

# Confidence interval method 2: Using lm 
# lm(Y ~ Tr, data = match$mdata)


```

Part 5
```{r}
lalonde['u74'] = lalonde$re74 == 0
lalonde['u75'] = lalonde$re75 == 0
glm1 = glm(treat ~ age + I(age^2) + educ + I(educ^2) + black + hisp + married + nodegr + re74 + I(re74^2) + re75 + I(re75^2) + u74 + u75, family = binomial, data = lalonde)

X = glm1$fitted.values
Y = lalonde$re78
Tr = lalonde$treat

rr = Match(Y = Y, Tr = Tr, X = X, M = 1)
summary(rr)

genout = GenMatch(Tr = Tr, X = X)
mout = Match(Y = Y, Tr = Tr, X = X, estimand = "ATE", Weight.matrix = genout)
summary(mout)

matched = lalonde[unlist(mout[c("index.treated","index.control")]), ]
matchedtab2 = CreateTableOne(vars = c("age","educ","black","hisp","married","nodegr","re74","re75"), strata ="treat", data = matched, test = FALSE)
print(matchedtab2, smd = TRUE)

# Genetic Matching via propensity scores
# genout_prop = GenMatch(Tr=nswre74wcps_data$treat, X=Xs_prop, BalanceMatrix=propscores,
#                        estimand="ATT", M=1, max.generations=10)
# mout_genmatch_prop = Match(Y=nswre74wcps_data$re78, Tr=nswre74wcps_data$treat,
#                       X=Xs_prop, estimand="ATT", M=1, Weight.matrix=genout_prop)
# summary(mout_genmatch_prop)
##estimate: 1614.6
##standard error: 936.87
##p value: 0.0848
##confidence interval: -221.666, 3450.865

# Genetic Matching using all covariates (first-order)
# genout = GenMatch(Tr=nswre74wcps_data$treat, X=Xs, BalanceMatrix=Xs,
#                        estimand="ATT", M=1, max.generations=10)
# mout_genmatch = Match(Y=nswre74wcps_data$re78, Tr=nswre74wcps_data$treat,
#                       X=Xs, estimand="ATT", M=1, Weight.matrix=genout)
# summary(mout_genmatch)
##estimate: 2066.7
##standard error:986.2
##p value: 0.03612
##confidence interval: 133.748, 3999.652

```



Include analysis of differing caliper scores
Include analysis of M
Include analysis of pop size
Include analysis of estiamnd
Include analysis of exact
Include analysis of replace

