---
title: "Quiz 2"
output: html_notebook
---

# Quiz 2 -- More R Coding...

## This quiz covers competencies you should know by the first day of class.
## The first ten questions are intended to be basic.
## The final five questions are more advanced.
## An answer key will be provided by Tuesday, July 24.

## NOTE: Most students find it helpful to code in R using R-Studio.
## This involves installing R, and installing R-Studio, and then 
## running R-Studio. Google it if you run into problems, and 
## if Google can't answer your questions, feel free to email me.

# Load the Matching library, as described at the beginning of the prior quiz
# (Quiz 1). If you've never installed this Matching library before,
# you'll have to begin by doing so -- one way is to use use "install.packages()",
# but there are also 'point-and-click' ways to do it.
# So---the first step is to install the library, and the next step is to load it.
```{r}
library(Matching)
```

# The "Matching" library comes with a data set called 
# GerberGreenImai which relates to a series of published papers.
# You can see some of those papers if (after loading the library)
# you type:
```{r}
?GerberGreenImai
```


# This command opens the help page for the data set. At the bottom of the help page,
# you'll see a list of 4 relevant published papers.

# To load the data set into working memory, type:
```{r}
data(GerberGreenImai)
```

# To take a quick look at the data set, type:
```{r}
head(GerberGreenImai)
```

# Question 1 (using "dim")
What are the dimensions of the data set (how many rows and columns)?
```{r}
dim(GerberGreenImai)
```

# Question 2 (using "names")
# What are the variable names?
```{r}
names(GerberGreenImai)
```


# Question 3 (using "is.na")
# Is there any missing data? If so, how many elements are missing 
# and which rows contain the missing data?
```{r}
any(is.na(GerberGreenImai))
which(is.na(GerberGreenImai), arr.ind = TRUE)
```


# Question 4 (using the "mode" function, or "is.factor")
# In this data set, each column is a variable (e.g., "AGE", "WARD", etc.). 
# Which of these variables (if any) is a factor? 
```{r}
names(GerberGreenImai)[sapply(GerberGreenImai, is.factor)]
```


# Question 5 (using the "mode" function)
# Code a for loop that outputs column name and variable type, per the below format:

# column name #1  variable type #1
# column name #2  variable type #2
# column name #3  variable type #3
# etc., etc., etc.,  type of each 
```{r}
for (i in (1:ncol(GerberGreenImai))) { print( cbind(paste('column name:', colnames(GerberGreenImai)[i]),paste('variable type:',sapply(GerberGreenImai, class)[i]))) }
```

# Question 6
# What do you think the variable AGE2 represents? 
```{r}
## Using 
## ?GerberGreenImai and checking function arguments gives
## Age Squared/100
```


# Question 7
# Which individuals (and how many) voted in 1978?
```{r}
which(GerberGreenImai["VOTED98"] == 1)
sum(GerberGreenImai["VOTED98"])
```


# Question 8
# Which individuals (and how many) have age > 35 and did NOT vote in 1978?
```{r}
which(GerberGreenImai["VOTED98"] == 0 & GerberGreenImai["AGE"] > 35)
nrow(GerberGreenImai[GerberGreenImai["VOTED98"] == 0 & GerberGreenImai["AGE"] > 35,])
```


# Question 9
# What line of code eliminates all rows that satisfy the conditions of Question 8?
```{r}
GerberGreenImai[!(GerberGreenImai["VOTED98"] == 0 & GerberGreenImai["AGE"] > 35),]
```


# Question 10
# Append a column that records a "1" if Question 8's conditions satisfy, 0 if not.
# Congrats, you just created a dummy variable! Google it if you have questions.
```{r}
GerberGreenImai['DummyCol'] = as.integer((GerberGreenImai["VOTED98"] == 0 & GerberGreenImai["AGE"] > 35))
```


### More advanced questions
# Question 11
# Write a function that performs the algorithm you identified in Question 6.
```{r}
all((GerberGreenImai$AGE ^2)/100 == GerberGreenImai$AGE2)
```


# Question 12
# Run a regression w/ AGE2 and AGE as dependent and independent vars (respectively).
# Do you get statistically significant results? Is there something causal going on?
# Does the model make predictions that are not possible in the real world? Explain.
```{r}
summary(lm(AGE2 ~ AGE, data= GerberGreenImai))
plot(AGE2 ~ AGE, data = GerberGreenImai)
```
Makes a linear model that stretches on forever - there is a limit to where the model makes sense.

# Question 13 (hint: see the help page for "formula")
# Consider the following code:
reg1 <- lm(PHN.C1 ~ PERSONS + VOTE96.1 + NEW + MAJORPTY + AGE + 
         WARD + I(PERSONS*VOTE96.1), data=GerberGreenImai[c(1:1000),])
# you can then type something like "summary(reg1)"...
                 
# This regression predicts whether or not a subject received a get-out-the-vote
# phone call based on the first 1000 observations. 

# What is meant by I(PERSONS*VOTE96.1)? Use the resulting regression
# model to make a prediction for individuals in rows 1001 and 1002. Do these 
# predictions make sense? Why or why not?
```{r}
summary(lm(PHN.C1 ~ PERSONS, data=GerberGreenImai[c(1:1000),]))
#summary(PHN.C1 ~ PERSONS + VOTE96.1 + NEW + MAJORPTY + AGE + WARD + I(PERSONS*VOTE96.1), data=GerberGreenImai[c(1:1000),])
```


# Question 14
# Consider the "WARD" variable. Observe the first 20 elements.
# What happens when you perform arithmetic or statistical operations on these
# elements? Why? Provide intuition for the result you obtain.
```{r}
GerberGreenImai$WARD[1:20]
GerberGreenImai$WARD[1]
GerberGreenImai$WARD[2]
GerberGreenImai$WARD[1] + GerberGreenImai$WARD[2]
```
Not meaningful for factors because Ward 13 + Ward 24 != Ward 37

# Question 15
# Write code that takes the "WARD" variable and creates dummy variables for
# each individual ward. We can call this "dummy-fying" a categorical variable.


# Given your answer to Question 14, do you see why you would want to "dummy-fy" 
# this variable prior to performing statistical analysis on it (and
# variables likes this one)?
```{r}
model.matrix(~GerberGreenImai$WARD)
```

