---
title: "R Notebook"
output: html_notebook
---


Import and clean data
```{r}
df = read.csv("Downloads/Investors day data - eventbrite export 041118 .csv", stringsAsFactors = FALSE)
investors = df[df$Ticket.Type == "Investor", c(19, 20, 21, 22, 25)]
startups = df[df$Ticket.Type == "Startup", c(9, 11, 12, 13, 14, 15, 16, 17, 18)]

# select only startups which want investment
startups = startups[grep("Getting investment", startups$What.are.you.interested.in.), ]
```

Create 2 matrices for rank-ordering the meeting criteria for startups and investors
```{r}
investorfields = strsplit(as.character(investors$What.is.your.search.field.),split='|', fixed=TRUE)
startupfields = strsplit(as.character(startups$What.is.your.field.), split = "|", fixed = TRUE)

startupfulfil = matrix(nrow = length(startupfields), ncol = length(investorfields))
investorfulfil = matrix(nrow = length(startupfields), ncol = length(investorfields))

for(i in 1:length(startupfields)){
  for(j in 1:length(investorfields)){
    startupfulfil[i,j] = length(intersect(startupfields[[i]], investorfields[[j]]))/length(startupfields[[i]])
    investorfulfil[i,j] = length(intersect(startupfields[[i]], investorfields[[j]]))/length(investorfields[[j]])
  }
}
```

Match based on each person getting 5 matches (optimization problem)
```{r}
A = matrix(nrow = nrow(investors), ncol = 2)
A[,1] = as.character(investors[, 5])

for (i in 1:nrow(investors)){
  # this is already the multivariate case. when given to others to experiment, what needs to be tweaked is the range of values for which it is acceptable. It is not in percentages, but in each of the variables' percentile.
  x = startups[sort(startupfulfil[,i], decreasing = TRUE, index.return = TRUE)$ix[1:5], 1]
  y = startups[sort(investorfulfil[,i], decreasing = TRUE, index.return = TRUE)$ix[1:8], 1]
  A[i,2] = paste(intersect(x,y), collapse = " | ")
}

A

```

```{r}
x = as.character(startups[sort(startupfulfil[,3], decreasing = TRUE, index.return = TRUE)$ix[1:10],2])
x[4:5] = c("Warsaw", "London")

dist = gmapsdistance(origin = c("Berlin"), 
              destination = c("Potsdam", "Telaviv", "Skopje", "Warsaw", "London", "Berlin", "Berlin", "Berlin", "Prishtina", "St.+Louis"), 
              mode = "driving",
              shape = "long")$Distance[,3]

for (i in strsplit(A[3,2], split = " | ", fixed = TRUE)[[1]]){
  print(gmapsdistance(origin = "Berlin", destination = as.character(startups[startups$Startup.Name. == i, 2]), mode = "driving")$Distance)
}
sort(dist, decreasing = FALSE, index.return = TRUE, na.last = TRUE)$ix

```

Create a function that takes in an investor, the number of matches required, (keywords), and a vector of their preferences. 
The preferences can be across - Industry, Stage of Product, Location, (Year), 
```{r}

matcher = function(matches = 5, preferences = c(75, 50, 20), keywords = FALSE){
  
  investorfields = strsplit(as.character(investors$What.is.your.search.field.),split='|', fixed=TRUE)
  startupfields = strsplit(as.character(startups$What.is.your.field.), split = "|", fixed = TRUE)
  startupfulfil = matrix(nrow = length(startupfields), ncol = length(investorfields))

  for(i in 1:length(startupfields)){for(j in 1:length(investorfields)){
      startupfulfil[i,j] = length(intersect(startupfields[[i]], investorfields[[j]]))/length(startupfields[[i]])}}}
  startupfulfil
  
  
matcher()

```

```{r}
newinvestors = read.csv("Downloads/Clean Investors Day Data - Investors.csv", stringsAsFactors = FALSE)
investorfields = strsplit(newinvestors$Field.of.interest, split = "|", fixed = TRUE)
startupfields = strsplit(as.character(startups$What.is.your.field.), split = "|", fixed = TRUE)
startupfulfil = matrix(nrow = length(startupfields), ncol = length(investorfields))

  for(i in 1:length(startupfields))
    {for(j in 1:length(investorfields)){
        startupfulfil[i,j] = length(intersect(startupfields[[i]], investorfields[[j]]))/length(startupfields[[i]])}}

listofrankings = list()

for(i in 1:length(investorfields)){
  x = matrix(unlist(sort(scale(startupfulfil)[,i], decreasing = TRUE, index.return = 1)), nrow = nrow(startups))
  x = data.frame(x[,1], startups[x[,2], c(1,3,5,8)]); names(x)[2] = paste0(newinvestors[i,1], " matches")
  listofrankings[[i]] = x
}
for (i in 1:length(listofrankings)){
  View(listofrankings[[i]])
}


```

